### 2016-07-29 AnalysizeStop.Rmd
072916

---
title: "Analyze Stop Data"
author: "Mengyu Li"
date: "July 22, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(lubridate)

# Read data
stop_by <- read_csv("cleaned_stop_data.csv")

stop_by$Call.Date.Time <- mdy_hm(stop_by$Call.Date.Time)
stop_by$AgeRange <- as.factor(stop_by$AgeRange)

stop_by <- stop_by %>%
  mutate(hour = hour(Call.Date.Time))

#Remove input "E" from the R column

stop_race <- ggplot(stop_by, aes(x = Reason, fill = Enforcement)) + 
  geom_bar(aes(y = ..count../sum(..count..))) + 
  labs(y = "Density") +
  facet_wrap(~Race)

stop_race <- ggplot(stop_by, aes(x = Reason, fill = Enforcement)) + 
  geom_bar(aes(y = ..count../sum(..count..))) + 
  labs(y = "Density") +
  scale_fill_brewer(palette = "Set1") +
  facet_grid(Race~Gender)

stop_race

View(stop_by)
```