### 2016-07-31 AnalysizeStop.Rmd

---
title: "Analyze Stop Data"
author: "Mengyu Li"
date: "July 22, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(lubridate)
library(stringr)
library(ggvis)

```

```{r}
# Read data
stop_by <- read_csv("cleaned_stop_data.csv")

stop_by$Call.Date.Time <- mdy_hm(stop_by$Call.Date.Time)
stop_by$AgeRange <- as.factor(stop_by$AgeRange)

stop_by <- stop_by %>%
  mutate(Hour = as.integer(hour(Call.Date.Time))) %>%
  mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))

```

```{r}
#Remove input "E" from the R column
stop_race <- stop_by %>%
  ggvis(x = ~Race , fill = ~AgeRange) %>%
  layer_bars() %>%
  add_tooltip(function(stop_by) stop_by$AgeRange) 
stop_race

```

```{r}
stop_day <- stop_by %>%
  ggvis(x = ~Day, fill = ~Race) %>%
  layer_bars() %>%
  add_tooltip(function(stop_by) stop_by$Race)
stop_day
```

```{r}
stop_time <- stop_by %>%
  ggvis(x = ~Hour, fill := "lightpink") %>%
  layer_histograms(width = input_slider(1, 12, value = 1, step = 2, label = "width"))
stop_time
#width = input_slider(1, 12, step = 2, label = "width")
```

```{r}

stop_time_percentage <- ggplot(stop_by, aes(x = Hour, fill = AgeRange)) + 
  geom_bar(aes(y = ..count../sum(..count..))) +
  labs(y = "Density") +
  facet_wrap(~AgeRange)
stop_time_percentage

```

```{r}
stop_by_hourCount <- stop_by %>%
  group_by(Hour) %>%
  summarise(hourcount = n())
stop_by_ageRangeCount <- stop_by %>%
  group_by(Hour, AgeRange) %>%
  summarise(ageCount = n())
stop_by_percentage <- left_join(stop_by_ageRangeCount, stop_by_hourCount)
stop_by_percentage <- stop_by_percentage %>%
  mutate(percentage = ageCount/hourcount)

stop_time_percentage2 <- ggplot(stop_by_percentage, aes(x = Hour, fill = AgeRange)) + 
  geom_bar(aes(y = percentage), stat="identity") +
  labs(y = "Density") +
  facet_wrap(~AgeRange)
stop_time_percentage2
```

```{r}
stop_by_hourCount <- stop_by %>%
  group_by(Hour) %>%
  summarise(hourcount = n())
stop_by_raceRangeCount <- stop_by %>%
  group_by(Hour, Race) %>%
  summarise(raceCount = n())
stop_by_percentage1 <- left_join(stop_by_raceRangeCount, stop_by_hourCount)
stop_by_percentage1 <- stop_by_percentage1 %>%
  mutate(percentage = raceCount/hourcount)

stop_time_percentage3 <- ggplot(stop_by_percentage1, aes(x = Hour, fill = Race)) + 
  geom_bar(aes(y = percentage), stat="identity") +
  labs(y = "Density") +
  facet_wrap(~Race)
stop_time_percentage3
```

```{r}
stop_by_reason_race <- stop_by %>%
  group_by(Race, Reason) %>%
  summarise(totalCount = n())
stop_by_arrest <- stop_by %>%
  filter(Enforcement == "A") %>%
  group_by(Race, Reason) %>%
  summarise(arrestCount = n())
stop_by_percentage_race <- left_join(stop_by_reason_race, stop_by_arrest)
stop_by_percentage_race <- stop_by_percentage_race %>%
  mutate(percentage = arrestCount / totalCount)

stop_arrest <- ggplot(stop_by_percentage_race, aes(x = Race, fill = Race)) + 
  geom_bar(aes(y = percentage), stat="identity") +
  labs(y = "Density") +
  facet_wrap(~Reason)
stop_arrest
```

```{r}
stop_by_reason_age <- stop_by %>%
  group_by(AgeRange, Reason) %>%
  summarise(totalCount = n())
stop_by_arrest_age <- stop_by %>%
  filter(Enforcement == "A") %>%
  group_by(AgeRange, Reason) %>%
  summarise(arrestCount = n())
stop_by_percentage_age <- left_join(stop_by_reason_age, stop_by_arrest_age)
stop_by_percentage_age <- stop_by_percentage_age %>%
  mutate(percentage = arrestCount / totalCount)

stop_arrest_age <- ggplot(stop_by_percentage_age, aes(x = AgeRange, fill = AgeRange)) + 
  geom_bar(aes(y = percentage), stat="identity") +
  labs(y = "Density") +
  facet_wrap(~Reason)
stop_arrest_age
```

```{r}
stop_by_reason_gender <- stop_by %>%
  group_by(Gender, Reason) %>%
  summarise(totalCount = n())
stop_by_arrest_gender <- stop_by %>%
  filter(Enforcement == "A") %>%
  group_by(Gender, Reason) %>%
  summarise(arrestCount = n())
stop_by_percentage_gender <- left_join(stop_by_reason_gender, stop_by_arrest_gender)
stop_by_percentage_gender <- stop_by_percentage_gender %>%
  mutate(percentage = arrestCount / totalCount)

stop_arrest_gender <- ggplot(stop_by_percentage_gender, aes(x = Gender, fill = Gender)) + 
  geom_bar(aes(y = percentage), stat="identity") +
  labs(y = "Density") +
  facet_wrap(~Reason)
stop_arrest_gender
```

```{r}
stop_race <- ggplot(stop_by, aes(x = Reason, fill = Enforcement)) + 
  geom_bar(aes(y = ..count../sum(..count..))) + 
  labs(y = "Density") +
  facet_wrap(~Race)
stop_race

```

```{r}

stop_race2 <- ggplot(stop_by, aes(x = Reason, fill = Enforcement)) + 
  geom_bar(aes(y = ..count../sum(..count..))) + 
  labs(y = "Density") +
  scale_fill_brewer(palette = "Set1") +
  facet_grid(Race~Gender)

stop_race2

```