### 2016-07-31 Stop_Data_Analysis.Rmd

---
title: "Stop Data Analysis"
author: "Mengyu Li"
date: "July 29, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### About the Berkeley PD Stop Data

The Berkeley PD stop data reflects accidents that have been visited by the Berkeley Police Department.