### 2016-08-02 final_paper/Shinyapp.Rmd

---
title: "SHINY APP"
author: "Shangjun Jiang"
date: "2 August 2016"
output: html_document
runtime: shiny
---

```{r}
setwd("~/Desktop/Stats R/PROJECT/finalproject_angry_ladies/final_paper")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sp)
library(ggmap)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(data.table)
library(rworldmap)
library(maps)
library(mapdata)
library(maptools)
library(scales)
library(RgoogleMaps)
library(tmap)
library(sp)
library(rgdal)
library(rgeos)
library(RColorBrewer)

```

```{r}
## Read in the Data:
#finaldf <- readRDS( 'StopData_finaldf.rds' )
mergedf <- readRDS( 'StopData_merged.rds' )
```

```{r}
## Location info:
latmax <- max( mergedf$lat, na.rm = TRUE ) 
latmin <- min( mergedf$lat, na.rm = TRUE )
lonmax <- max( mergedf$long, na.rm = TRUE ) 
lonmin <- min( mergedf$long, na.rm = TRUE )
latvals <- c( latmin, latmax )
lonvals <- c( lonmin, lonmax )
```

```{r}
## Make the variables factors:
mergedf$CarSearch <- factor(mergedf$CarSearch)
levels(mergedf$CarSearch) <- c("No Search", "Search")

mergedf$Race <- factor(mergedf$Race)
levels(mergedf$Race) <- c("Asian", "Black", "Hispanic", "Other", "White")

mergedf$Enforcement <- factor(mergedf$Enforcement)
levels(mergedf$Enforcement) <- c("Arrest", "Citation", "Other", "Warning")

mergedf$Reason <- factor(mergedf$Reason)
levels(mergedf$Reason) <- c("Other", "Investigation", "Probation/Parole", "Reasonable Suscipcion", "Traffic", "Wanted")

mergedf <- mergedf %>%
  mutate(Arrest = ifelse( 
    as.character(Enforcement) == "Arrest", 
    "Arrested",
    "Not Arrested") ) %>%
  mutate(Arrest = factor(Arrest))

v <- factor(mergedf$Other)
levels(v)

mergedf <- mergedf %>%
  mutate(Emergency.Psych.Eval = ifelse( str_detect(as.character(Other), "MH"), 
                                        yes = "Yes",
                                        no = "No") )

mergedf$Emergency.Psych.Eval <- factor(mergedf$Emergency.Psych.Eval)

?write.csv

write.csv(mergedf, "mergedf.csv", row.names = F)

berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14)
```

## MAP 4:
```{r}

ggmap(berkMap) +
  geom_point(aes(x = long, y = lat, colour = Reason), data = mergedf, alpha = 0.7, size = 3) +
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + ggtitle("All BPD Stops, by Reason 2015-2016") +
    facet_wrap( ~ Reason , ncol = 2)
```