### 2016-08-04 final_paper/Some thoughts.Rmd

---
title: "Some ideas"
author: "Shangjun Jiang"
date: "4 August 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sp)
library(ggmap)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(data.table)
library(rworldmap)
library(maps)
library(mapdata)
library(maptools)
library(scales)
library(RgoogleMaps)
library(tmap)
library(sp)
library(rgdal)
library(rgeos)
library(RColorBrewer)
setwd("~/Desktop/Stats R/PROJECT/finalproject_angry_ladies/final_paper")
```

## Data Loading

```{r, echo=FALSE}
# Read data

############################## stop data ############################## 
stop_by <- read_csv("cleaned_stop_data.csv")

stop_by$Call.Date.Time <- mdy_hm(stop_by$Call.Date.Time)
stop_by$AgeRange <- as.factor(stop_by$AgeRange)

stop_by <- stop_by %>%
  mutate(Hour = as.integer(hour(Call.Date.Time))) %>%
  mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))

stop_by$Race<- stop_by$Race %>% 
  str_replace("A", "Asian") %>% 
  str_replace("B", "Black") %>% 
  str_replace("H", "Hispanic") %>% 
  str_replace("O", "Other") %>% 
  str_replace("W", "White")

stop_by$Gender<- stop_by$Gender %>% 
  str_replace("F", "Female") %>% 
  str_replace("M", "Male")

stop_by$Reason <- stop_by$Reason %>% 
  str_replace("I", "Investigation") %>% 
  str_replace("T", "Traffic") %>% 
  str_replace("R", "Reasonable Suspicion") %>% 
  str_replace("R", "Probation/ Parole") %>% 
  str_replace("W", "Wanted")

stop_by$Enforcement <- stop_by$Enforcement %>% 
  str_replace("A", "Arrest") %>% 
  str_replace("C", "Citation") %>% 
  str_replace("O", "Other") %>% 
  str_replace("W", "Warning")

stop_by$CarSearch <- stop_by$CarSearch %>% 
  str_replace("S", "Search") %>% 
  str_replace("N", "No Search")

############################## location data ############################## 
mergedf <- readRDS( 'StopData_merged.rds' )

## Location info:
latmax <- max(mergedf$lat, na.rm = TRUE) 
latmin <- min(mergedf$lat, na.rm = TRUE)
lonmax <- max(mergedf$long, na.rm = TRUE) 
lonmin <- min(mergedf$long, na.rm = TRUE)
latvals <- c(latmin, latmax)
lonvals <- c(lonmin, lonmax)

## Make the variables factors:
mergedf$CarSearch <- factor(mergedf$CarSearch)
levels(mergedf$CarSearch) <- c("No Search", "Search")

mergedf$Race <- factor(mergedf$Race)
levels(mergedf$Race) <- c("Asian", "Black", "Hispanic", "Other", "White")

mergedf$Enforcement <- factor(mergedf$Enforcement)
levels(mergedf$Enforcement) <- c("Arrest", "Citation", "Other", "Warning")

mergedf$Reason <- factor(mergedf$Reason)
levels(mergedf$Reason) <- c("Other", "Investigation", "Probation/Parole", "Reasonable Suscipcion", "Traffic", "Wanted")

mergedf <- mergedf %>%
  mutate(Arrest = ifelse( 
    as.character(Enforcement) == "Arrest", 
    "Arrested",
    "Not Arrested") ) %>%
  mutate(Arrest = factor(Arrest))

v <- factor(mergedf$Other)

mergedf <- mergedf %>%
  mutate(Emergency.Psych.Eval = ifelse( str_detect(as.character(Other), "MH"), 
                                        yes = "Yes",
                                        no = "No") )

mergedf$Emergency.Psych.Eval <- factor(mergedf$Emergency.Psych.Eval)

### BerkeleyMap 
berkMap = map = get_map(location = c(lon = mean(lonvals), lat = mean(latvals)), zoom = 14)

############################## census data Rebecca ##############################

blocks <- readOGR("Census_Tract_Polygons2010","Census_tracts_2010", verbose = TRUE)

locationCensusFiles <- "Census_Tract_Polygons2010"
blocks <- readOGR(locationCensusFiles,"Census_tracts_2010", verbose = TRUE)
b2 <- spTransform(blocks, CRS("+proj=longlat +datum=WGS84"))
b3 <- fortify(b2)

############################## census data JENNY ############################## 
census <- read.csv("Census_Data_2000_And_2010.csv")
census2010<- census[census$Year==2010,]

racecensus2010 <- census2010[census2010$Heading=="Not Hispanic or Latino"|census2010$Heading=="HISPANIC OR LATINO AND RACE",]
sexcensus2010 <- census2010[census2010$Heading=="Sex",]
agecensus2010 <- census2010[census2010$Heading=="Age",]

race <- select(racecensus2010, Description, Amount) 
race$Description <- race$Description %>% 
  str_replace_all("Hispanic or Latino.*", "Hispanic") %>% 
  str_replace("Black or African American", "Black") %>% 
  str_replace("American Indian and Alaska Native", "Other") %>% 
  str_replace("Native Hawaiian and Other Pacific Islander", "Other") %>% 
  str_replace("Some other race", "Other") %>% 
  str_replace("Two or more races", "Other")

race <- race[race$Description !="Not Hispanic",]

race <- race %>% 
  group_by(Description) %>% 
  tally(Amount)

names(race) <- c("Description", "Counts")

sex <- select(sexcensus2010, Description, Amount)

age <- select(agecensus2010, Description, Amount) 
age <- age[age$Description != "Median age",]
age$Description <- c("0-18", "0-18", "18-64", "65+")

age<- age %>% 
  group_by(Description) %>% 
  tally(Amount)
names(age) <- c("Description", "Counts")

############################## arrest data ############################## 
arrest <- read.csv("Berkeley_PD_Log_-_Arrests.csv")

############################## jail data ############################## 
jail <- read.csv("Berkeley_PD_Log_-_Jail_Bookings.csv")

############################## call for service data ##############################

callservice<- read.csv("Berkeley_PD_-_Calls_for_Service.csv")

callservice<- callservice %>% 
  mutate(latlong= 
           str_replace_all(callservice$Block_Location,"[0-9]* [A-Za-z]*", "") %>%
           str_replace_all("\nBerkeley,\n", "") %>% 
           str_replace("[A-Za-z]*", "") %>% 
           str_replace(";","") %>% 
           str_replace("&","") %>% 
           str_replace("[A-Za-z]*", ""))

callservice$latlong<- str_replace(callservice$latlong, "\\(", "") %>% 
                    str_replace("\\)", "")

callservice<- callservice %>% 
  separate(latlong, c("lat", "long"), sep=",")

callservice$lat <- as.numeric(callservice$lat)
callservice$long <- as.numeric(callservice$long)

############################## multiplot function ############################## 
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```

## PART I: Racial Discrimination??

```{r}
p1 <- ggplot(stop_by)+
  geom_bar(aes(x=Race),
           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))+
  labs(title = "Stop data by race",
       x = "Race",
       y = "Counts")

p2 <- ggplot(race)+
  geom_bar(aes(x=Description, 
               y=Counts),
           stat = "identity", 
           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))+
  labs(title = "Demographic race data",
       x = "Race",
       y = "Counts")

multiplot(p1, p2,cols=2)

```

#### arrest + jail data

## PART II: time of being stopped

#### by race
#### by age

#### location of being stopped
#### census data + stop data map + call for service map

## PART IV: Exploration
#### pick location + time
#### time you top 3 stop reasons + top 2 race + top 2 age groups