### 2016-08-07 eda/Heatmap/HeatmapExploration4.Rmd
heatmaps

---
title: "Heat Map Exploration"
author: "Rebecca Reus"
date: "August 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r, include = FALSE, echo = FALSE, warning = FALSE}
library(sp)
library(ggmap)
library(tidyr)
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(maps)
library(mapdata)
library(rgdal)
library(rgeos)
#library(gpclib)       # loads polygon clipping library
library(maptools)     # loads sp library too
library(RColorBrewer) # creates nice color schemes
library(classInt)     # finds class intervals for continuous variables
library(gdata)
library(XLConnect)
library(data.table)
library(gridExtra)
```

### Set data file names:
```{r}
# data file names:
data_file <- "../../clean_data/StopData_merged2.rds" # use the cleaned version of StopData_merged.rds for plotting. See StopData_clean_merged_rds.R for more information.
census_tract_folder <- "../../raw_data/Census_Tract_Polygons2010"
census_tract_shp <- "Census_tracts_2010"
census_tract_csv <- "../../raw_data/Census_tracts_2010.csv"
alameda_xls <- "../../raw_data/2010_Pop_Block_County.xls"
alameda_tract_nums_xls <- "../../raw_data/census_tract_numbers.xls"
alameda_clean_csv <- "../../raw_data/alameda_clean_census.csv"
berkeley_clean_csv <- "../../clean_data/berkeley_census_clean.csv"
```

### Read in the data:
```{r}
# read in the data to modify:
df <- readRDS( data_file )
berkpop <- read.csv2(berkeley_clean_csv)
berkpop <- berkpop[,-1]
a3 <- read.csv2(alameda_clean_csv)
a3 <- a3[,-1]
```

```{r, eval = F, include = F}
# Used to clean the alameda data, do not run.
wb <- loadWorkbook(alameda_xls)
alameda <- readWorksheet(wb, sheet=1,startRow = 5) 
```

```{r}
# Location info (to set lim values for coordinates):
latmax <- max( df$lat, na.rm = TRUE )
latmin <- min( df$lat, na.rm = TRUE )
lonmax <- max( df$long, na.rm = TRUE )
lonmin <- min( df$long, na.rm = TRUE )
latvals <- c( latmin, latmax )
lonvals <- c( lonmin, lonmax )
```

### Get the regular Berkeley map:
```{r,include = F, echo = FALSE}
# get rid of the axes:
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )

# get the plain Berkeley map from Google:
berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 12)

berkgg <-ggmap(berkMap) +
  xlim(-122.335, lonmax) + ylim(latvals) +
  ditch_the_axes 
berkgg
```

### Get the census shape files:
```{r, include=FALSE, echo = FALSE}

census <- readOGR(census_tract_folder,census_tract_shp, verbose = TRUE)

censusT <- spTransform(census, CRS("+proj=longlat +datum=WGS84"))
censusT@data$ID <- as.numeric(censusT@data$ID)

censusCSV <- read.csv(file=census_tract_csv, header = TRUE)
```

### Examine total population by census block:
```{r, include = F, message = F}
# READ IN DATA RIGHT!!!
census.df <- fortify(censusT, region = 'GEOID10')
pop <- read.csv(file=census_tract_csv, header = TRUE)
names(pop)[13] <- "id"
pop <- pop %>%
  select(GEOID10, id, TotalPop) %>% 
  mutate( id = str_c("0", as.character(id)) ) 
census.df_pop <- left_join(census.df, pop, by = "id")
```

```{r, include = F}
popdenmap <- berkgg + 
  geom_polygon(data = census.df_pop, aes(x = long, y = lat, group = group, fill = TotalPop), color = "white", alpha = .7) 
popdenmap
```

```{r,include = F}
pd2 <- popdenmap + 
    scale_fill_gradientn(colours = rev(rainbow(7) ) ) 
pd2
```

## Population density maps:
```{r}
grid.arrange(popdenmap,pd2, ncol = 2)
```

```{r, include = F, echo = F}
bpop2 <- berkpop %>%
  select(-Total.population..not.Hispanic.or.Latino) %>%
   select(-Two.or.More.Races) %>%
  select(-One.race.total)

bpop2$Other = bpop2$American.Indian.and.Alaska.Native +
                  bpop2$Native.Hawaiian.and.Other.Pacific.Islander +
                  bpop2$Some.Other.Race

bpop2 <- bpop2 %>%
  mutate(Percent.Black = Black.or.African.American/Total.population) %>%
  mutate(Percent.Other = Other/Total.population) %>%
  mutate(Percent.White = White/Total.population) %>%
  mutate(Percent.Asian = Asian/Total.population) %>%
  mutate(Percent.Hispanic = Hispanic.or.Latino/Total.population) %>%
  mutate(Percent.Berkeley = Total.population/sum(Total.population)) %>%
  select(-TotalPop) %>%
   select(-American.Indian.and.Alaska.Native,-Native.Hawaiian.and.Other.Pacific.Islander, -Some.Other.Race)

# %>%
#   mutate(Relative.Per.Black = Black.or.African.American/sum(Black.or.African.American)) %>%
#   mutate(Relative.Per.Other = Other/sum(Other)) %>%
#   mutate(Relative.Per.White = White/sum(White)) %>%
#   mutate(Realtive.Per.Asian = Asian/sum(Asian)) %>%
#   mutate(Relative.Per.Hispanic = Hispanic.or.Latino/sum(Hispanic.or.Latino)) %>%

names(bpop2)[2] <- "Population"
names(bpop2)[3] <- "Hispanic"
names(bpop2)[5] <- "Black"
names(bpop2)

# write the clean berkeley census 2010 data to file:
#write_csv(x = bpop2, path = "../../clean_data/berk_census2010_clean.csv")

```

```{r, include = FALSE, echo = FALSE, warning=FALSE}
# read in the data:
census.df <- fortify(censusT, region = 'NAME10')
census.df <- left_join(census.df, bpop2, by = "id")
#write_csv(x = census.df, path = "../../clean_data/berk_census2010_mapdata.csv")
```

Population by Race:
```{r, include = F, echo = F, warning = FALSE}

# create color palette:

p <- colorRampPalette(c("white", "red"))(128)
palette(p)

# black population:

black1 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Black), color = "white", alpha = .5)  +
  ggtitle("Berkeley Census 2010:\nBlack Population") + 
  scale_fill_gradientn(colors = p) +
  labs(fill = "Count")

black2 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Percent.Black), color = "white", alpha = .7) +
  ggtitle("Berkeley Census 2010:\nPercent Black") +
  scale_fill_gradientn(colors = p) +
  labs(fill = "Density")

# white population:

w1 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = White), color = "white", alpha = .5)  +
  ggtitle("Berkeley Census 2010:\nWhite Population") + 
  scale_fill_gradientn(colors = p) +
  labs(fill = "Count")

w2 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Percent.White), color = "white", alpha = .7) +
  ggtitle("Berkeley Census 2010:\nPercent White") +
  scale_fill_gradientn(colors = p) +
  labs(fill = "Density")

# asian population:

a1 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Asian), color = "white", alpha = .5)  +
  ggtitle("Berkeley Census 2010:\nAsian Population") + 
  scale_fill_gradientn(colors = p) +
  labs(fill = "Count")

a2 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Percent.Asian), color = "white", alpha = .7) +
  ggtitle("Berkeley Census 2010:\nPercent Asian") +
  scale_fill_gradientn(colors = p) +
  labs(fill = "Density")

# hispanic population:

h1 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Hispanic), color = "white", alpha = .5)  +
  ggtitle("Berkeley Census 2010:\nHispanic Population") + 
  scale_fill_gradientn(colors = p) +
  labs(fill = "Count")

h2 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Percent.Hispanic), color = "white", alpha = .7) +
  ggtitle("Berkeley Census 2010:\nPercent Hispanic") +
  scale_fill_gradientn(colors = p) +
  labs(fill = "Density")

# other population:

o1 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Other), color = "white", alpha = .5)  +
  ggtitle("Berkeley Census 2010:\nOther Population") + 
  scale_fill_gradientn(colors = p) +
  labs(fill = "Count")

o2 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Percent.Other), color = "white", alpha = .7) +
  ggtitle("Berkeley Census 2010:\nPercent Other") +
  scale_fill_gradientn(colors = p) +
  labs(fill = "Density")

```

View side by side:
```{r, fig.width = 15}
grid.arrange(w1, black1, h1, a1, o1, ncol = 5)
grid.arrange(w2, black2, h2, a2, o2, ncol = 5)
```

### BPD Stops:
```{r}
# edit bpd stops:
# nblacks <- sum(df$Race == "Black")
# nasians <- sum(df$Race == "Asian")
# nwhite <- sum(df$Race == "White")
# nother <- sum(df$Race == "Other")
# nhisp <- sum(df$Race == "Hispanic")
# totalpop <- sum(berkpop$TotalPop)
# stop.summary <- data.frame(Race = unique(df$Race))

stop.summary <- df %>%  
   group_by(Race) %>%
   summarise(Stopped=n() ) %>%
  mutate(Percent.Stopped = Stopped/nrow(df))

racepops <- c(sum(bpop2$Asian), sum(bpop2$Black), sum(bpop2$Hispanic), sum(bpop2$Other), sum(bpop2$White) )

stop.summary$Race.Population = racepops
stop.summary$Percent.Race.Stopped = stop.summary$Stopped/stop.summary$Race.Population

#write.csv(x = stop.summary, file = "../../clean_data/StopData_summary.csv")

```

```{r, include = F, echo = F, warning = F, message = F}
# all stop density:
allbpd_density <- berkgg +
  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha =  ..level../100),
                 size = 2, bins = 5, data = df, geom = "polygon") +
  scale_fill_gradientn( colors = p) +
  theme (panel.grid.major = element_blank (), # remove major grid
         panel.grid.minor = element_blank ()  # remove minor grid
  ) + 
  ggtitle("All BPD Stops Density, 2015-2016") +
  labs(fill = "Stops") +
  guides(alpha = FALSE)
allbpd_density

# by race:
bpd_race <- berkgg +
  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha =  ..level../100),
                 size = 2, bins = 5, data = df, geom = "polygon") +
  scale_fill_gradientn(trans = "log10", colors = p) +
  theme (panel.grid.major = element_blank (), # remove major grid
         panel.grid.minor = element_blank ()  # remove minor grid
  ) + 
  guides(alpha = FALSE) +
  facet_grid( ~ Race)
bpd_race
```

```{r, eval = T}
allbpd_density
bpd_race
```

```{r. eval = F}
# data file names:
data_file <- "../../raw_data/StopData_merged2.rds" # use the cleaned version of StopData_merged.rds for plotting. See StopData_clean_merged_rds.R for more information.
census_tract_folder <- "../../raw_data/Census_Tract_Polygons2010"
census_tract_shp <- "Census_tracts_2010"
census_tract_csv <- "../../raw_data/Census_tracts_2010.csv"
berkeley_clean_csv <- "../../raw_data/berkeley_census_clean.csv"
census <- readOGR(census_tract_folder,census_tract_shp, verbose = TRUE)

censusT <- spTransform(census, CRS("+proj=longlat +datum=WGS84"))
censusT@data$ID <- as.numeric(censusT@data$ID)

censusCSV <- read.csv(file=census_tract_csv, header = TRUE)
census.df <- fortify(censusT, region = 'NAME10')

#berkcensus2010map <- census.df

df <- readRDS( data_file )
berkpop <- read.csv2(berkeley_clean_csv)
berkpop <- berkpop[,-1]

bpop2 <- berkpop %>%
  select(-Total.population..not.Hispanic.or.Latino) %>%
   select(-Two.or.More.Races) %>%
  select(-One.race.total)

bpop2$Other = bpop2$American.Indian.and.Alaska.Native +
                  bpop2$Native.Hawaiian.and.Other.Pacific.Islander +
                  bpop2$Some.Other.Race

bpop2 <- bpop2 %>%
  mutate(Percent.Black = Black.or.African.American/Total.population) %>%
  mutate(Percent.Other = Other/Total.population) %>%
  mutate(Percent.White = White/Total.population) %>%
  mutate(Percent.Asian = Asian/Total.population) %>%
  mutate(Percent.Hispanic = Hispanic.or.Latino/Total.population) %>%
  mutate(Percent.Berkeley = Total.population/sum(Total.population)) %>%
  select(-TotalPop) %>%
   select(-American.Indian.and.Alaska.Native,-Native.Hawaiian.and.Other.Pacific.Islander, -Some.Other.Race)

names(bpop2)[2] <- "Population"
names(bpop2)[3] <- "Hispanic"
names(bpop2)[5] <- "Black"
names(bpop2)

#berkcensus2010$id <- as.character(berkcensus2010$id)
census.df <- left_join(census.df, bpop2, by = "id")
berkcensus2010map <- census.df

saveRDS(berkcensus2010map, "../../clean_data/berk_census2010_mapdata.rds")

```