### 2016-08-07 eda/Heatmap/Results.Rmd

---
title: "Results"
author: "Rebecca Reus"
date: "August 7, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, echo = FALSE, warning = FALSE}
library(sp)
library(ggmap)
library(tidyr)
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(maps)
library(mapdata)
library(rgdal)
library(rgeos)
#library(gpclib)       # loads polygon clipping library
library(maptools)     # loads sp library too
library(RColorBrewer) # creates nice color schemes
library(classInt)     # finds class intervals for continuous variables
library(gdata)
library(XLConnect)
library(data.table)
library(gridExtra)
library(pander)
library(formattable)

# data file names:

file_stops <- "../../clean_data/StopData_clean.csv" 
file_berkcensus2010 <- "../../clean_data/berk_census2010_clean.csv" 
file_berkcensus2010map <-"../../clean_data/berk_census2010_mapdata.rds" 
file_stopsummary <- "../../clean_data/StopData_summary.csv"

# read in the data:

stops <- read_csv(file = file_stops)
berkcensus2010 <- read_csv(file = file_berkcensus2010)
berkcensus2010map <- readRDS(file_berkcensus2010map)
stopsummary <- read.csv( file_stopsummary )
stopsummary <- select(stopsummary, -X)

# Location info (to set lim values for coordinates):

latmax <- max( stops$lat, na.rm = TRUE )
latmin <- min( stops$lat, na.rm = TRUE )
lonmax <- max( stops$long, na.rm = TRUE )
lonmin <- min( stops$long, na.rm = TRUE )
latvals <- c( latmin, latmax )
lonvals <- c( lonmin, lonmax )

# get rid of the axes theme:

ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )

# get the plain Berkeley map from Google:
berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 12)

berkgg <-ggmap(berkMap) +
  xlim(-122.335, lonmax) + ylim(latvals) +
  ditch_the_axes 
#berkgg

```

```{r, include = FALSE, echo = FALSE, warning = FALSE}

# create color palette:

p <- colorRampPalette(c("white", "red"))(128)
palette(p)

# create the maps:
popdenmap <- berkgg + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Berkeley), color = "white", alpha = .7) +
  ggtitle("2010 Population Density") +
  labs(fill = "Percent") +
  scale_fill_gradientn(colors = p)
popdenmap

# pd2 <- popdenmap + 
#   scale_fill_gradientn(colours = rev(rainbow(7) ) ) +
#   ggtitle("Berkeley Census 2010:\nPopulation")
# pd2

# black population:

black2 <- berkgg + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Black), color = "white", alpha = .7) +
  scale_fill_gradientn(colors = p, limits = c(0,1)) +
  guides( fill = "none" ) +
  ggtitle("% Black")

# white population:

w2 <- berkgg + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.White), color = "white", alpha = .7) +
  scale_fill_gradientn(colors = p, limits = c(0,1)) +
  guides( fill = "none" ) +
  ggtitle("% White")

# asian population:

a2 <- berkgg + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Asian), color = "white", alpha = .7) +
  scale_fill_gradientn(colors = p, limits = c(0,1)) +
  guides( fill = "none" ) +
  ggtitle("% Asian")

# hispanic population:

h2 <- berkgg + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Hispanic), color = "white", alpha = .7) +
  scale_fill_gradientn(colors = p, limits = c(0,1)) +
  guides( fill = "none" ) +
  ggtitle( "% Hispanic")

# other population:

o2 <- berkgg + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Other), color = "white", alpha = .7) +
  scale_fill_gradientn(colors = p, limits = c(0,1)) +
  guides( fill = "none" ) +
  ggtitle("% Other")
```

## Population Density Maps

```{r, warning = FALSE, echo = F, fig.width = 5, fig.height = 5}
popdenmap
```

```{r, warning = FALSE, echo = FALSE,  fig.width = 15, fig.height = 5}
grid.arrange(w2,black2,a2,h2,o2, ncol = 5)
```

## Comparing Stop Density to Population Density

```{r, message = F, echo = FALSE}
ss <- stopsummary %>%
  select(Race, Percent.Stopped, Percent.Race.Stopped)

ss$Percent.Stopped <- percent(ss$Percent.Stopped)
ss$Percent.Race.Stopped <- percent(ss$Percent.Race.Stopped)

names(ss)[2:3] <- c("Percent Stopped", "Percent Population Stopped")

panderOptions("digits", 2)
pander(ss)

bargraph1 <- ggplot(ss, aes(Race, `Percent Stopped`, fill = Race))+
  geom_bar(stat = "identity") +
  guides( fill = FALSE)
bargraph2 <- ggplot(ss, aes(Race, `Percent Population Stopped`, fill = Race))+
  geom_bar(stat = "identity") +
  guides( fill = FALSE)
grid.arrange(bargraph1, bargraph2, ncol = 2)
```