### 2016-08-08 eda/PaperDrafts/Some thoughts.Rmd

---
title: "Some ideas"
author: "Shangjun Jiang"
date: "4 August 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sp)
library(ggmap)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(data.table)
library(rworldmap)
library(maps)
library(mapdata)
library(maptools)
library(scales)
library(RgoogleMaps)
library(tmap)
library(sp)
library(rgdal)
library(rgeos)
library(RColorBrewer)
setwd("~/Desktop/Stats R/PROJECT/finalproject_angry_ladies/final_paper")
```

## Data Loading

```{r, echo=FALSE}
# Read data

############################## location data ############################## 
mergedf <- readRDS( 'StopData_merged.rds' )

## Location info:
latmax <- max(mergedf$lat, na.rm = TRUE) 
latmin <- min(mergedf$lat, na.rm = TRUE)
lonmax <- max(mergedf$long, na.rm = TRUE) 
lonmin <- min(mergedf$long, na.rm = TRUE)
latvals <- c(latmin, latmax)
lonvals <- c(lonmin, lonmax)

## Make the variables factors:
mergedf$CarSearch <- factor(mergedf$CarSearch)
levels(mergedf$CarSearch) <- c("No Search", "Search")

mergedf$Race <- factor(mergedf$Race)
levels(mergedf$Race) <- c("Asian", "Black", "Hispanic", "Other", "White")

mergedf$Enforcement <- factor(mergedf$Enforcement)
levels(mergedf$Enforcement) <- c("Arrest", "Citation", "Other", "Warning")

mergedf$Reason <- factor(mergedf$Reason)
levels(mergedf$Reason) <- c("Other", "Investigation", "Probation/Parole", "Reasonable Suscipcion", "Traffic", "Wanted")

mergedf <- mergedf %>%
  mutate(Arrest = ifelse( 
    as.character(Enforcement) == "Arrest", 
    "Arrested",
    "Not Arrested") ) %>%
  mutate(Arrest = factor(Arrest))

v <- factor(mergedf$Other)

mergedf <- mergedf %>%
  mutate(Emergency.Psych.Eval = ifelse( str_detect(as.character(Other), "MH"), 
                                        yes = "Yes",
                                        no = "No") )

mergedf$Emergency.Psych.Eval <- factor(mergedf$Emergency.Psych.Eval)

mergedf$Call.Date.Time <- mdy_hm(mergedf$Call.Date.Time)
mergedf$AgeRange <- as.factor(mergedf$AgeRange)

mergedf <- mergedf %>%
  mutate(Hour = as.integer(hour(Call.Date.Time))) %>%
  mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))

mergedf$Gender<- mergedf$Gender %>% 
  str_replace("F", "Female") %>% 
  str_replace("M", "Male")

mergedf$AgeRange<- mergedf$AgeRange %>% 
  str_replace("1", "0-18") %>% 
  str_replace("2", "18-29") %>% 
  str_replace("3", "30-39") %>%  
  str_replace("4", "40+")

### BerkeleyMap 
berkMap = map = get_map(location = c(lon = mean(lonvals), lat = mean(latvals)), zoom = 14)

############################## census data Rebecca ##############################

blocks <- readOGR("Census_Tract_Polygons2010","Census_tracts_2010", verbose = TRUE)

locationCensusFiles <- "Census_Tract_Polygons2010"
blocks <- readOGR(locationCensusFiles,"Census_tracts_2010", verbose = TRUE)
b2 <- spTransform(blocks, CRS("+proj=longlat +datum=WGS84"))
b3 <- fortify(b2)

############################## census data JENNY ############################## 
census <- read.csv("Census_Data_2000_And_2010.csv")
census2010<- census[census$Year==2010,]

racecensus2010 <- census2010[census2010$Heading=="Not Hispanic or Latino"|census2010$Heading=="HISPANIC OR LATINO AND RACE",]
sexcensus2010 <- census2010[census2010$Heading=="Sex",]
agecensus2010 <- census2010[census2010$Heading=="Age",]

race <- select(racecensus2010, Description, Amount) 
race$Description <- race$Description %>% 
  str_replace_all("Hispanic or Latino.*", "Hispanic") %>% 
  str_replace("Black or African American", "Black") %>% 
  str_replace("American Indian and Alaska Native", "Other") %>% 
  str_replace("Native Hawaiian and Other Pacific Islander", "Other") %>% 
  str_replace("Some other race", "Other") %>% 
  str_replace("Two or more races", "Other")

race <- race[race$Description !="Not Hispanic",]

race <- race %>% 
  group_by(Description) %>% 
  tally(Amount)

names(race) <- c("Description", "Counts")

sex <- select(sexcensus2010, Description, Amount)

age <- select(agecensus2010, Description, Amount) 
age <- age[age$Description != "Median age",]
age$Description <- c("0-18", "0-18", "18-64", "65+")

age<- age %>% 
  group_by(Description) %>% 
  tally(Amount)
names(age) <- c("Description", "Counts")

############################## arrest data ############################## 
arrest <- read.csv("Berkeley_PD_Log_-_Arrests.csv")

############################## jail data ############################## 
jail <- read.csv("Berkeley_PD_Log_-_Jail_Bookings.csv")

############################## call for service data ##############################

callservice<- read.csv("Berkeley_PD_-_Calls_for_Service.csv")

callservice<- callservice %>% 
  mutate(latlong= 
           str_replace_all(callservice$Block_Location,"[0-9]* [A-Za-z]*", "") %>%
           str_replace_all("\nBerkeley,\n", "") %>% 
           str_replace("[A-Za-z]*", "") %>% 
           str_replace(";","") %>% 
           str_replace("&","") %>% 
           str_replace("[A-Za-z]*", ""))

callservice$latlong<- str_replace(callservice$latlong, "\\(", "") %>% 
                    str_replace("\\)", "")

callservice<- callservice %>% 
  separate(latlong, c("lat", "long"), sep=",")

callservice$lat <- as.numeric(callservice$lat)
callservice$long <- as.numeric(callservice$long)

############################## multiplot function ############################## 
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```

## PART I: Racial Discrimination??

```{r, fig.height = 5, fig.width = 10, echo = FALSE}

m <- stop_by %>% 
  group_by(Race) %>% 
  tally()
names(m) <- c("Race", "stop")
names(race) <- c("Race", "census")

pop <- left_join(race,m,by = "Race")
pop <- pop %>% 
  mutate(percentage=(stop/census)) %>% 
pop$percentage<- round(pop$percentage, digits = 2)

ggplot(pop)+
  geom_bar(aes(x=Race,
               y=percentage), 
           stat = "identity",
           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40b8d0"))+
  geom_text(aes(x = Race, 
                y = percentage, 
                label = paste0(percentage,"%")), 
            size=5,
            colour= "#000000")+
  scale_y_continuous(labels=percent, limits = c(0,1)) +
  labs(title = "The percentage of being stopped by police (%)",
       x = "Race",
       y = "Percentage") +
  theme(axis.line = element_line(size=1, colour = "black"), 
        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
        panel.border = element_blank(), panel.background = element_blank()) + 
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"), 
        text=element_text(family="Tahoma"),
        axis.text.x=element_text(colour="black", size = 10), 
        axis.text.y=element_text(colour="black", size = 10))

```

```{r}
race %>%
  ggvis(x = ~Description, y = ~Counts) %>%
  layer_bars(fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF")) %>%
  add_tooltip(function(race) race$Counts) 
  labs(title = "Counts of people recorded in every day in a week")

```

#### arrest + jail data

## PART II: time of being stopped

#### by race
#### by age

#### location of being stopped
#### census data + stop data map + call for service map

## PART IV: Exploration
#### pick location + time
#### time you top 3 stop reasons + top 2 race + top 2 age groups